---
title: "Projet"
author: "moi"
date: "2024-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r }
library(ggplot2)
library(reshape2)
library(gridExtra)


```

```{r }
load('velib.RData')
summary(velib)


```

```{r }

# data preparation
loading = as.matrix(velib$data)
colnames(loading) = 1:ncol(loading)
rownames(loading) = velib$names

stations = 1:nrow(loading)
coord = velib$position[stations,]
coord$bonus = velib$bonus[stations]

# select exactly 7 days of data (we remove the first 13 dates)
dates = 14:181
loading = loading[stations, dates]
colnames(loading) = 1:length(dates)

coord <- coord[!duplicated(coord[,1]), ]



```


```{r }
head(loading)



```

```{r }
head(coord)



```

LOADING AT A STATION


```{r }

i = 1
options(
  repr.plot.width = 20,     # Largeur en pouces
  repr.plot.height = 10     # Hauteur en pouces
)
plot(loading[i,])
abline(v = seq(0,length(loading[i,]), by = 24), col = "red", lty = 1)

```

```{r }
options(repr.plot.width = 15, repr.plot.height = 10)

# --- #

timeTick = 1 + 24*(0:6)  # vector corresponding to the beginning of days

# select 16 stations
stations = sample.int(nrow(x), 16 , replace=TRUE)

df = melt(x[stations,])  #the function melt reshapes it from wide to long

p = list()
for (i in 1:16){
    dfi = df[df$Var1 == velib$names[stations[i]],]
    p[[i]] = ggplot(dfi, aes(x=Var2, y=value)) + 
        geom_line(col="darkorchid") + 
        geom_vline(xintercept=timeTick, col="orange", linetype="dashed") +
        labs(title=velib$names[stations[i]])
}
do.call(grid.arrange,p)

```



```{r }

str(coord)
```


BOXPLOT OF THE VARIABLE

```{r }
options(repr.plot.width = 15, repr.plot.height = 6)

boxplot(loading)

```

AVERAGE LOADING .....


```{r }
### TO BE COMPLETED ### 

print('--- Average fill rate ---')
mean(loading)

# --- #
print('')
means = list()
for (i in 1:length(loading[,1]))
    {
    means[i] = mean(loading[i,])
     }
indice_min = which.min(means)
indice_max = which.max(means)

print('--- Least crowded station, on average ---')
print(velib$names[indice_min])

# --- #
print('')

print('--- Fullest station, on average ---')
print(velib$names[indice_max])


```


Does the average load vary from one station to another?



```{r }



plot(seq(1, length(means)), means)
abline(h = mean(loading), col = "red", lty = 1)

```

Does the average load vary over the course of a day?

Plot the average hourly loading for each day (on a single graph).

```{r }


mean_per_hour_per_day = colMeans(loading)
mean_per_hour_per_day = matrix(mean_per_hour_per_day, nrow = 24)
mean_per_hour         = rowMeans(mean_per_hour_per_day)

# --- #

mean_per_hour_per_day            = as.data.frame(mean_per_hour_per_day)
colnames(mean_per_hour_per_day)  = list("Monday", "Tuesday", "Wednesday","Thursday", "Friday", "Saturday", "Sunday")
mean_per_hour_per_day$time_range = c(1:24)
mean_per_hour_per_day            = melt(mean_per_hour_per_day, id='time_range', variable.name='Days')

mean_per_hour            = as.data.frame(mean_per_hour)
colnames(mean_per_hour)  = list("Weekly")
mean_per_hour$time_range = c(1:24)

# --- #

options(repr.plot.width = 15, repr.plot.height = 10)

ggplot() + 
    geom_line(data=mean_per_hour_per_day, aes(x=time_range, y=value, color=Days)) + 
    geom_line(data=mean_per_hour, aes(x=time_range, y=Weekly), linewidth = 1.5)

```

ETUDE DESCRIPTIVE DE LOADING

```{r}

library(corrplot)


cor_matrix <- cor(loading, use = "complete.obs")


corrplot(cor_matrix, method = "color")


```


ACP

```{r}
library(FactoMineR)
acp <- PCA(loading, scale.unit = TRUE,
           graph = FALSE, quali.sup = 1, quanti.sup = 2, ncp = 7)
# Décroissance des valeurs propres
library(factoextra)
g1<-fviz_eig(acp, addlabels = TRUE, ylim = c(0, 40))
library(reshape2)
g2<-ggplot(melt(acp$ind$coord),aes(x=Var2,y=value))+
  geom_boxplot()+
  xlab("")
grid.arrange(g1,g2,ncol=2)




```

```{r}

fviz_pca_var(acp)
fviz_pca_ind(acp,col.ind="contrib",label="none",gradient.cols = c("white", "#2E9FDF", "#FC4E07" ))
fviz_pca_var(acp,axes=c(1,3))
fviz_pca_ind(acp,col.ind="contrib",label="none",gradient.cols = c("white", "#2E9FDF", "#FC4E07" ),axes=c(1,3))
```

K MEANS 

```{r}

km.loading <- kmeans(loading, centers = 2)

# Représentation dans les coordonnées de l'acp
acp2 <- PCA(cbind(clus = as.factor(km.loading$cluster),
          loading), scale.unit = TRUE,
          graph = FALSE, quali.sup = 1:2, ncp = 7)


fviz_pca_ind(acp2, label="none", habillage="clus")

```


















LDA

```{r}

str(loading)

```

```{r}

head(loading)


```


```{r}


```


```{r}


```

```{r}


```


































Velib Station Map



```{r }



library(ggmap)
```

VELIB STATION MAP

Where are the velib stations located?

    Plot the stations coordinates on a 2D map (latitude vs. longitude)
    Use the average hourly loading as a color scale
    You can consider different times of day, for example 6am, 12pm, 11pm on Monday, or the average weekly load at 6am.
    You can consider different days at the same time, or the average load for each day.
    You can use the qmplot function of the ggmap to charge the map of Paris


```{r }
ggmap::register_stadiamaps(key = "f5428ead-0f2c-4508-bd37-50d807a861ca")

## TO BE COMPLETED ### 
## Simple 2D representation
# Monday at hour 6h, 12h, 23h

# Hours to be displayed
h     = 18
hours = seq(h, 168, 24)
load_per_hour = rowMeans(loading[,hours])

df = coord
df$loading = load_per_hour

# --- #

options(repr.plot.width = 10, repr.plot.height = 10)

qmplot(data=df, longitude, latitude, color=loading) +
    labs(title = paste('Stations loading - Weekly average at',h,'h'))

```

```{r }

## Simple 2D representation
# Loading at 6pm, depending on the day of the week

options(repr.plot.width = 20, repr.plot.height = 15)

h = 18
hours = seq(h, 168, 24)
days  = list("Monday", "Tuesday", "Wednesday","Thursday", "Friday", "Saturday", "Sunday")

dfi = coord
p = list()
for (i in 1:7){
    dfi$loading = loading[,hours[i]]
    p[[i]] = ggplot(dfi, aes(x=longitude, y=latitude, color=loading)) + 
        geom_point() +
        labs(title = paste("Stations loading - ", days[i], h,"h"))
}

do.call(grid.arrange,c(p, ncol=3))
```

```{r }

h     = 18
hours = seq(h, 168, 24)
load_per_hour = rowMeans(loading[,hours])

df = coord
df$loading = load_per_hour

# --- #

options(repr.plot.width = 10, repr.plot.height = 10)

qmplot(data=df, longitude, latitude, color=loading) +
    labs(title = paste('Stations loading - Weekly average at',h,'h'))

```



Influence of Altitude Difference on Station Loading



```{r}


df = data.frame(size=c(sum(coord$bonus==0), sum(coord$bonus==1)),
                labels = c('No hill','Hill'))

ggplot(df, aes(x="", y=size, fill=labels)) +
    geom_bar(stat="identity", width=1) +
    geom_text(aes(label=size), position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    scale_color_hue(direction = -1) +  # to reverse the default colormap order
    theme_void()




```





```{r }
options(repr.plot.width = 10, repr.plot.height = 10)

qmplot(data=coord, longitude, latitude, color=hill) +
    scale_color_hue(direction = -1) +
    labs(title = 'Hilltop stations')

```

```{r }

coord$hill = as.factor(coord$bonus)
levels(coord$hill) = c('No hill','Hill')

# --- #

options(repr.plot.width = 10, repr.plot.height = 10)

ggplot(coord, aes(x=longitude, y=latitude, color=hill)) + 
    geom_point() +
    scale_color_hue(direction = -1) +
    labs(title = 'Hilltop stations')
```



ETUDE DESCRIPTIVE COORD


```{r }

library(corrplot)


cor_matrix <- cor(coord[,c(1,2)], use = "complete.obs")


corrplot(cor_matrix, method = "color")

```

```{r }


```

```{r }

library(FactoMineR)
acp <- PCA(coord[,c(1,2)], scale.unit = TRUE,
           graph = FALSE, quali.sup = 1, quanti.sup = 2, ncp = 7)
# Décroissance des valeurs propres
library(factoextra)
g1<-fviz_eig(acp, addlabels = TRUE, ylim = c(0, 40))
library(reshape2)
g2<-ggplot(melt(acp$ind$coord),aes(x=Var2,y=value))+
  geom_boxplot()+
  xlab("")
grid.arrange(g1,g2,ncol=2)


```

```{r }
head(coord[,c(1,2)])

```

```{r }

str(loading)

```

```{r }
str(coord)

```
PCA sur Loading
```{r }

pca_result <- PCA(loading, scale.unit = TRUE, ncp = min(150, ncol(loading)), graph = FALSE)
plot(pca_result$eig[,2], type = "b", main = "Variance Explained by PCA Components", xlab = "Principal Component", ylab = "Percentage of Variance Explained")


```

```{r }
fviz_pca_ind(pca_result,
             label = "none",  # This prevents the labels from being shown
             col.ind = "cos2",  # Color by the quality of representation on the dimensions (optional)
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),  # Color gradient (optional)
             repel = TRUE  # Avoid overplotting (optional)
)

```

```{r }
str(coord)

```

```{r }
# Perform PCA
pca_result <- prcomp(coord[, -3], scale. = TRUE)  # Exclude 'bonus' column

# Determine the number of principal components to retain
summary(pca_result)

# Plot the proportion of variance explained by each principal component
plot(pca_result, type = "l")

# Cumulative proportion of variance explained
cumsum(pca_result$sdev^2) / sum(pca_result$sdev^2)


```

```{r }


```

```{r }


```

```{r }


```

```{r }


```
